### git {.unnumbered}

To use **git** to version your projects, you need to store your credentials 
(username and email) on your laptop.

Open `Positron`and run these lines in the {{< fa brands r-project >}} console:

```{r}
#| echo: true
#| eval: false

# Install 'gert' R package ----
install.packages("gert")

# Store your git user name ----
gert::git_config_global_set(
  name = "user.name", 
  value = "Jane Doe"            ## replace this value
)

# Store your git user email ----
gert::git_config_global_set(
  name = "user.email", 
  value = "jane.doe@mail.com"   ## replace this value
)
```

When you create a new `git` project, the default branch is named **_master_**. 
It's recommended to replace **_master_** by **_main_** (for more information read this
[article](https://www.zdnet.com/article/github-to-replace-master-with-alternative-term-to-avoid-slavery-references/)).

Run this line in the {{< fa brands r-project >}} console:

```{r}
#| echo: true
#| eval: false

# Rename git default branch name ----
gert::git_config_global_set(
  name = "init.defaultBranch", 
  value = "main"
)
```

From now, when you will initialize `git` in a project, the default branch 
will be **main**.

Check your `git` configuration:

```{r}
#| echo: true
#| eval: false

# Print git global configuration ----
gert::git_config_global()
#               name                 value      level
#          user.name              Jane Doe     global
#         user.email     jane.doe@mail.com     global
# init.defaultbranch                  main     global
```
