### GitHub API token {.unnumbered}

If you want to use the GitHub API, essentially with the 
{{< fa brands r-project >}} packages
[`usethis`](https://usethis.r-lib.org/) and 
[`rcompendium`](https://cran.r-project.org/package=rcompendium),
you need to create a 
[GitHub Personal Access Token](https://docs.github.com/fr/authentication/keeping-your-account-and-data-secure/creating-a-personal-access-token) (PAT).

Proceed as follow:

1. Go to [https://github.com/settings/tokens](https://github.com/settings/tokens)
2. Click on **Generate new token (classic)**
3. Choose a name for your new token (e.g. your laptop name)
4. Select no expiration date
5. In the section _Select scopes_, select at least **repo** and **workflow**
6. Copy this token

Then open `Positron`and run these lines in the {{< fa brands r-project >}} console:

```{r}
#| echo: true
#| eval: false

## Install usethis package ----
install.packages("usethis")

## Open ~/.Renviron file ----
usethis::edit_r_environ()
# • Modify '~/.Renviron'
# • Restart R for changes to take effect
```

The last command will open the `~/.Renviron` file. Add this new line: 
`GITHUB_PAT='ghp_XXX'` where `ghp_XXX` is your token.
Do not forget to restart the R session.


<!--

Finally let's tell `usethis` to use the SSH protocol to communicate with GitHub. 
Open **RStudio** and run:

```{r}
#| echo: true
#| eval: false

## Open ~/.Rprofile file ----
usethis::edit_r_profile()
# • Modify '~/.Rprofile'
# • Restart R for changes to take effect
```

In the `~/.Rprofile` file, add this new line: `options(usethis.protocol = "ssh")`.
Do not forget to restart **RStudio Desktop**.

-->

Finally, let's check that your GitHub PAT is accessible from {{< fa brands r-project >}}:

```{r}
#| echo: true
#| eval: false

Sys.getenv("GITHUB_PAT")
# ghp_XXX
```

::: {.callout-note}
## To go further

- The `.Renviron` file is a good place to store secrets (API tokens, password, etc.).
This file is read by {{< fa brands r-project >}} at startup. To access these variables,
you can use the function `Sys.getenv("GITHUB_PAT")`.

- The `.Rprofile` file allows you to run {{< fa brands r-project >}} command lines 
at startup. You can customize the default behaviour of {{< fa brands r-project >}}
and define some variables with `options()`. To access these variables,
you can use `options()$"usethis.protocol"`.

More information [here](https://cran.r-project.org/web/packages/startup/vignettes/startup-intro.html).
:::
