# macOS {.unnumbered}


## Installation

| Software        | Description                                    | Website                                                                    |
| :-------------- | :--------------------------------------------- | :------------------------------------------------------------------------: |
| R               | The R environment                              | [link](https://cran.r-project.org/)                                        |
| XQuartz         | Graphical window system                        | [link](https://www.xquartz.org/)                                           |
| RStudio Desktop | Integrated development environment (IDE) for R | [link](https://posit.co/download/rstudio-desktop/)                         |
| Pandoc          | Document converter used by `rmarkdown`         | [link](https://pandoc.org/)                                                |
| Quarto CLI      | Scientific publishing system built on Pandoc   | [link](https://quarto.org/)                                                |
| Git             | Version control software                       | [link](https://git-scm.com/)                                               |
| Docker Desktop  | Containerization software                      | [link](https://docs.docker.com/desktop/)                                   |


We are going to install the required software by using 
[**Homebrew**](https://brew.sh), a package manager for macOS. 
A package manager is a way to get software (and software updates) onto your 
machine without much work on your part. It's the macOS equivalent of `yum`, 
`pacman` or `apt-get`.

Before installing **Homebrew**, we need to install the Apple software
**Xcode Command Line Tools**. It is a self-contained package for software developers 
who wish to build Mac apps using UNIX-style commands.

Open a [Terminal](https://support.apple.com/en-gb/guide/terminal/apd5265185d-f365-44cb-8b09-71a064a42125/mac), 
run this line and accept the license:

```sh
## Install Xcode Command Line Tools ----
sudo xcode-select --install
```

We can now install **Homebrew** itself by running:

```sh
## Install Homebrew ----
/bin/bash -c "$(curl -fsSL https://raw.githubusercontent.com/Homebrew/install/HEAD/install.sh)"

## Check installation ----
brew --version
# Homebrew 4.1.19

## Update repositories (list of available software) ----
brew update
```

It's time to install the required softwares:

```sh
## Install utilities for R ----
brew install --cask xquartz     ## Graphical window system
brew install gfortran           ## FORTRAN compiler for GCC

## Install R ----
brew install --cask r

## Install RStudio Desktop ----
brew install --cask rstudio

## Install git ----
brew install git

## Install pandoc ----
brew install pandoc

## Install Quarto ----
brew install --cask quarto

## Install Docker ----
brew install docker
```


In addition, we need to install the {{< fa brands r-project >}} package 
[`quarto`](https://quarto-dev.github.io/quarto-r/):

```{r}
#| echo: true
#| eval: false

## Install quarto package ----
install.packages("quarto")
```


Finally, let's install a `LaTeX` distribution to convert documents from `.Rmd` 
(or `.qmd`) to `.pdf`. 
We are going to install a lightened distribution w/ the {{< fa brands r-project >}}
package [`tinytex`](https://yihui.org/tinytex/).


```{r}
#| echo: true
#| eval: false

## Install tinytex package ----
install.packages("tinytex")

## Install LaTeX distribution ----
tinytex::install_tinytex()
```


::: {.callout-important}
## Important

If you already have a `LaTeX` distribution, do not use `tinytex`.
:::





## Check

{{< fa exclamation-triangle >}} &nbsp; **Restart your machine**.

Open **RStudio** and run:

```{r}
#| echo: true
#| eval: false

## Get R version ----
R.version.string
# "R version 4.3.2 (2023-10-31)"

## Check if git is installed ----
Sys.which("git")
#            git
# "/usr/bin/git"

## Check if Rtools is installed ----
Sys.which("make")
#            make 
# "/usr/bin/make"

## Check if Pandoc is installed ----
Sys.which("pandoc")
#                  pandoc 
# "/usr/local/bin/pandoc"

## Check if Quarto is installed ----
Sys.which("Quarto")
#                  Quarto 
# "/usr/local/bin/Quarto"

## Check if LaTeX is installed ----
Sys.which("pdflatex")
#                  pdflatex 
# "/usr/local/bin/pdflatex" 

## Check if Docker is installed ----
Sys.which("docker")
#                  docker 
# "/usr/local/bin/docker" 

## Install package from sources ----
install.packages("jsonlite", type = "source")
```




::: {.callout-note}
## Update your system regularly

Keep your macOS Apps up-to-date by running:

```sh
## Check for updates ----
brew update

## Update macOS Apps ----
brew upgrade --greedy
```
:::



## Configuration

You've just installed a working environment for data science w/ 
{{< fa brands r-project >}}, {{< fa brands github >}}, and {{< fa brands docker >}}.
<br/>
**Now follow instruction to [configure your system](rstudio.html)**.
