# Ubuntu/Debian {.unnumbered}

## Installation

| Software        | Description                                    | Website                                                                    |
| :-------------- | :--------------------------------------------- | :------------------------------------------------------------------------: |
| R               | The R environment                              | [link](https://cran.r-project.org/)                                        |
| RStudio Desktop | Integrated development environment (IDE) for R | [link](https://posit.co/download/rstudio-desktop/)                         |
| Pandoc          | Document converter used by `rmarkdown`         | [link](https://pandoc.org/)                                                |
| Quarto CLI      | Scientific publishing system built on Pandoc   | [link](https://quarto.org/)                                                |
| Git             | Version control software                       | [link](https://git-scm.com/)                                               |
| Docker          | Containerization software                      | [link](https://docs.docker.com/desktop/)                                   |


We are going to install the required software by using 
[**APT**](https://en.wikipedia.org/wiki/APT_(software)), a package manager for 
Debian derived Linux distributions. A package manager is a way to get software 
(and software updates) onto your machine without much work on your part.

First check your **version of Ubuntu** in a terminal:


```sh
## Check Ubuntu version ----
cat /etc/issue
# Ubuntu 22.04 LTS
```


| Release   | Release name    | Repository URL                                               |
|:---------:|:----------------|:-------------------------------------------------------------|
| 22.04 LTS | Jammy Jellyfish | https://cloud.r-project.org/bin/linux/ubuntu jammy-cran40/   |
| 20.04 LTS | Focal Fossa     | https://cloud.r-project.org/bin/linux/ubuntu focal-cran40/   |
| 18.04 LTS | Bionic Beaver   | https://cloud.r-project.org/bin/linux/ubuntu bionic-cran40/  |
: Table 1. List of CRAN repositories for different versions of Ubuntu


Let's install {{< fa brands r-project >}} by following these steps 
(change the URL below by the one specific to your version of Ubuntu, cf. Table 1) :

```sh
## Install APT utilities ----
sudo apt install build-essential software-properties-common wget

## Add CRAN GPG key ----
wget -qO- https://cloud.r-project.org/bin/linux/ubuntu/marutter_pubkey.asc | sudo tee -a /etc/apt/trusted.gpg.d/cran_ubuntu_key.asc

## Add CRAN repository to APT repositories list ----
sudo add-apt-repository 'deb https://cloud.r-project.org/bin/linux/ubuntu jammy-cran40/'

## Update packages list ----
sudo apt update

## Install R ----
sudo apt install r-base r-base-dev
```


To install **RStudio Desktop**, follow these steps 
(uncomment the appropriate line matching your Ubuntu version):


```sh
## Download RStudio Desktop installer (Ubuntu = 22.04) ----
# wget https://download1.rstudio.org/electron/jammy/amd64/rstudio-2023.09.1-494-amd64.deb

## Download RStudio Desktop installer (Ubuntu = 20.04) ----
# wget https://download1.rstudio.org/electron/focal/amd64/rstudio-2023.09.1-494-amd64.deb

## Install RStudio Desktop ----
sudo dpkg -i rstudio-2023.09.1-494-amd64.deb

## Remove installer ----
rm rstudio-2023.09.1-494-amd64.deb
```

Now let's install **git**:

```sh
## Install git -----
sudo apt install git
```

And **Pandoc** (to convert documents with `rmarkdown`):

```sh
## Install pandoc -----
sudo apt install pandoc
```


Let's install the **Quarto** software, a multi-language, 
next-generation version of R Markdown developed by Posit (formerly RStudio Inc.), that 
includes dozens of new features and capabilities.


```sh
## Download Quarto installer ----
wget https://github.com/quarto-dev/quarto-cli/releases/download/v1.3.450/quarto-1.3.450-linux-amd64.deb

## Install Quarto ----
sudo dpkg -i quarto-1.3.450-linux-amd64.deb

## Remove installer ----
rm quarto-1.3.450-linux-amd64.deb
```

Finally, let's install **Docker Desktop**:

```sh
## Add Docker's official GPG key ----
sudo apt install ca-certificates curl gnupg
sudo install -m 0755 -d /etc/apt/keyrings
curl -fsSL https://download.docker.com/linux/ubuntu/gpg | sudo gpg --dearmor -o /etc/apt/keyrings/docker.gpg
sudo chmod a+r /etc/apt/keyrings/docker.gpg

## Add the repository to Apt sources ----
echo \
  "deb [arch="$(dpkg --print-architecture)" signed-by=/etc/apt/keyrings/docker.gpg] https://download.docker.com/linux/ubuntu \
  "$(. /etc/os-release && echo "$VERSION_CODENAME")" stable" | \
  sudo tee /etc/apt/sources.list.d/docker.list > /dev/null
  
sudo apt update

## Install Docker ----
sudo apt install docker-ce

## Allow user to use Docker without sudo ----
sudo usermod -aG docker ${USER}
```

In addition, we need to install the {{< fa brands r-project >}} package 
[`quarto`](https://quarto-dev.github.io/quarto-r/):

```{r}
#| echo: true
#| eval: false

## Install quarto package ----
install.packages("quarto")
```


Finally, let's install a `LaTeX` distribution to convert documents from `.Rmd` 
(or `.qmd`) to `.pdf`. 
We are going to install a lightened distribution w/ the {{< fa brands r-project >}}
package [`tinytex`](https://yihui.org/tinytex/).


```{r}
#| echo: true
#| eval: false

## Install tinytex package ----
install.packages("tinytex")

## Install LaTeX distribution ----
tinytex::install_tinytex()
```


::: {.callout-important}
## Important

If you already have a `LaTeX` distribution, do not use `tinytex`.
:::





## Check

{{< fa exclamation-triangle >}} &nbsp; **Restart your machine**.

Open **RStudio** and run:

```{r}
#| echo: true
#| eval: false

## Get R version ----
R.version.string
# "R version 4.3.2 (2023-10-31)"

## Check if git is installed ----
Sys.which("git")
#            git
# "/usr/bin/git"

## Check if make is installed ----
Sys.which("make")
#            make 
# "/usr/bin/make"

## Check if Pandoc is installed ----
Sys.which("pandoc")
#            pandoc 
# "/usr/bin/pandoc"

## Check if Quarto is installed ----
Sys.which("Quarto")
#            Quarto 
# "/usr/bin/Quarto"

## Check if LaTeX is installed ----
Sys.which("pdflatex")
#            pdflatex 
# "/usr/bin/pdflatex" 

## Check if Docker is installed ----
Sys.which("docker")
#            docker 
# "/usr/bin/docker" 

## Install package from sources ----
install.packages("jsonlite", type = "source")
```




::: {.callout-note}
## Update your system regularly

Keep your Linux packages up-to-date by running:

```sh
## Check for updates ----
sudo apt update

## Update Linux Apps ----
sudo apt upgrade
```
:::



## Configuration

You've just installed a working environment for data science w/ 
{{< fa brands r-project >}}, {{< fa brands github >}}, and {{< fa brands docker >}}.
<br/>
**Now follow instruction to [configure your system](rstudio.html)**.
